---
title: "Prediction_Assignment_Writeup"
author: "Franklin Santos"
date: "8/31/2020"
output: html_document
---

### Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

### Overview
The goal of this project is to predict the manner in which they did the exercise. This is the “classe” variable in the training set. This report describes how data was cleaned, how I split “pml-training.csv” into train set and test set, and some of models are investigated.

### Data Processing and Results
1. Loading add-on package and set seed
```{r}
set.seed(12345)
library(caret)
library(randomForest)
```
2. Download rawdata and submit_data
```{r}
url_train <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
rawdata <- read.csv(url_train, na.strings = c("", "NA"))
url_submit <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
submit_data <- read.csv(url_submit, na.strings = c("", "NA"))
```

3. Cleaning data

We should delete the column that contains NA to avoid the error. In addition, in order to make accurate predictions, columns that is not related exercise must also be deleted. In particular “X”, “user_name”, “raw_timestamp_part_1”, “raw_timestamp_part_2”, “cvtd_timestamp”, “new_window”, “num_window” are deleted.

```{r}
#Remove NA cols
colname <- colnames(rawdata)[!colSums(is.na(rawdata)) > 0]
colname

#Remove NA cols from submit data
colnamesub <- colnames(submit_data)[!colSums(is.na(rawdata)) > 0]
colnamesub

#Slice data related with exercise
colname <- colname[8: length(colname)]
df_wo_NA <- rawdata[colname]

#Submit data related with exercise
colnamesub <- colnamesub[8: length(colnamesub)]
submit_NA <- submit_data[colnamesub]


#Check the colnames of df_wo_NA is in submit_data.
#The last colname is "classe"
is.element(colname, colnames(submit_data))

df_wo_NA$classe <- factor(df_wo_NA$classe)
```

4. Split data into random train and test
```{r}
inTrain = createDataPartition(df_wo_NA$classe, p = 3/4)[[1]]
training = df_wo_NA[ inTrain,]
testing = df_wo_NA[-inTrain,]

#Other option for model_rf
training.ids <- createDataPartition(df_wo_NA$classe, p = 0.7, list = FALSE)
```

5. Random Forest

It takes a very long time for training, but it has a high accuracy.
```{r}
model_rf <- randomForest(x = df_wo_NA[training.ids, 1:52],
                         y = df_wo_NA[training.ids, 53],
                         ntree = 500,
                         keep.forest = TRUE)

pred_rf <- predict(model_rf, testing)
confusionMatrix(testing$classe, pred_rf)
```

6. Liner Discriminant Analysis

It takes a short time but poor accuracy.

```{r}
model_lda <- train(classe ~ ., data = training, method = "lda")
pred_lda <- predict(model_lda, testing)
confusionMatrix(testing$classe, pred_lda)
```

7. Recursive Partitioning and Regression Trees

The results can be confirmed visually, but poor accuracy.

```{r}
model_rpart <- train(classe ~ ., data = training, method = "rpart")
pred_rpart<- predict(model_rpart, testing)
confusionMatrix(testing$classe, pred_rpart)
```

```{r}
library(rattle)
fancyRpartPlot(model_rpart$finalModel)
```

8. Submit data with Random Forest

We can use the high accuracy model to submit data. In this report the Random Forest accuracy has the highest value 99.92. We can show the prediction.

```{r}
submit_rf <- predict(model_rf, submit_NA)
submit_rf
```

